Particle-in-cell simulation of ion and electron acceleration in trans-relativistic shocks.

### 2017-11-11 papers/Physica2017/biblio.tex
paper

\section*{References}
\begin{thebibliography}{20}
\bibitem{Bell1978} Bell A R 1978 \textit{MNRAS} \textbf{182} 147
\bibitem{Blandford1978} Blandford R D and Ostriker J P 1978 \textit{Astrophys. J.} \textbf{221} L29 
\bibitem{Berezhko2003} Berezhko E G, Ksenofontov L T and V{\"o}lk H J  2003 \textit{A}{\&}\textit{A} \textbf{412} L11
\bibitem{Uchiyama2007} Uchiyama Y, Aharonian F A, Tanaka T, Takahashi T and Maeda Y 2007 \textit{Nature} \textbf{449} 576
\bibitem{Lapenta2006} Lapenta G, Brackbill J U and Ricci P 2006 \textit{Phys. Plasmas} \textbf{13} 055904
\bibitem{Noguchi2007} Noguchi K, Tronci C, Zuccaro G and Lapenta G 2007 \textit{Phys. Plasmas} \textbf{14} 042308
\bibitem{Weibel1959} Weibel E S 1959 \textit{Phys. Rev. Lett.} \textbf{2} 83
\bibitem{Yoon1987} Yoon P H and Davidson R C 1987 \textit{Phys. Rev. A} \textbf{35} 2718
\bibitem{Park2015} Park J, Caprioli D and Spitkovsky A 2015 \textit{Phys. Rev. Lett.} \textbf{114} 085003
\end{thebibliography}

>>>>

\section*{References}
\begin{thebibliography}{20}
\bibitem{Bell1978} Bell A R 1978 \textit{MNRAS} \textbf{182} 147
\bibitem{Blandford1978} Blandford R D and Ostriker J P 1978 \textit{ApJ} \textbf{221} L29 
\bibitem{Bykov2014} Bykov A M, Ellison D C, Osipov S M and Vladimirov A E 2014 \textit{ApJ} \textbf{789} Issue 2, 137
\bibitem{Berezhko2003} Berezhko E G, Ksenofontov L T and V{\"o}lk H J  2003 \textit{A}{\&}\textit{A} \textbf{412} L11
\bibitem{Uchiyama2007} Uchiyama Y, Aharonian F A, Tanaka T, Takahashi T and Maeda Y 2007 \textit{Nature} \textbf{449} 576
\bibitem{Romansky2016} Romansky V I, Bykov A M, Osipov S M and Gladilin P E \textit{Journal of Physics: CS}
\bibitem{Lapenta2006} Lapenta G, Brackbill J U and Ricci P 2006 \textit{Phys. Plasmas} \textbf{13} 055904
\bibitem{Noguchi2007} Noguchi K, Tronci C, Zuccaro G and Lapenta G 2007 \textit{Phys. Plasmas} \textbf{14} 042308
\bibitem{Ellison2013} Ellison D C, Warren D C and Bykov A M 2013 \textit{ApJ} \textbf{776} Issue 1, 46
\bibitem{Sironi2011} Sironi L and Spitkovsky A 2011 \textit{ApJ} \textbf{741} Issue 1, 39
\end{thebibliography}

### 2017-11-11 papers/Physica2017/shock.tex
paper

\section{Shock wave simulation}
We studied the influence of fractions of heavy ions on the shock wave, in particular, the spectrum of particles and the shape of the shock wave. We present results of two simulations with different compositions of the plasma : pure protons and electrons in the first case and with the admixture of alpha particles in the second case. Initially the homogeneous flux is  moving from the right free boundary to the left. On the left side there is a reflecting super-conducting wall, which causes formation of shock wave. Simulations are one-dimensional and have following parameters: initial velocity $v = 0.2c$, number densities $n_e = 10^{-4} \rm{cm}^{-3}$, $n_p = 10^{-4} \rm{cm}^{-3}$ in the first simulation and $n_e = 10^{-4} \rm{cm}^{-3}$, $n_p = 0.6\cdot10^{-4} \rm{cm}^{-3}$, $n_\alpha = 0.2\cdot10^{-4} \rm{cm}^{-3}$ in the second, temperature $5\cdot10^8 \rm{K}$, magnetic field $B_\parallel = 10^{-4} \rm{G}, B_\perp= 0.7\cdot10^{-4} \rm{G}$, the full size of the box $L = 1\cdot10^{12} \rm{cm}$, the number of cells $N=2\cdot10^4$. Electron mass is reduced to $m_e = \frac{m_p}{20}$. The full time of simulation is $T = 2000 {\omega_p}^{-1}$.
\begin{figure}[h!]
	\centering
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/protons.eps} 
		\caption{Distribution of protons in the simulation without alpha particles.}
		\label{protons}
	\end{minipage}\hfill
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/protons_with_He.eps} 
		\caption{Distribution of protons in the simulation with alpha particles.}
		\label{protons_with_alpha}
	\end{minipage}
\end{figure}
\begin{figure}[h!]
	\centering
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/electrons.eps} 
		\caption{Distribution of electrons in the simulation without alpha particles.}
		\label{electrons}
	\end{minipage}\hfill
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{fig/electrons_with_He.eps} 
		\caption{Distribution of electrons in the simulation with alpha particles.}
		\label{electrons_with_alpha}
	\end{minipage}
\end{figure}

The results presented in Figures \ref{protons}-\ref{electrons_with_alpha} show that the values of $F(p)p^4$ for high energies are much greater for protons, than for electrons in both simulations. Also electrons need much more time to be accelerated and to form spectrum $\propto p^{-4}$. It means that protons are injected into the acceleration process more efficiently, and it is consistent with the work of Park et al. {\cite{Park2015}}. Also we have shown that minority of heavy ions increases the spectrum of protons and do not have influence on the spectrum of electrons. It can be explained by the fact, that heavy ions form the large scale turbulence and protons can efficiently scatter on this turbulence. Otherwise for electrons turbulence produced by protons is already enough large-scale  and the influence of heavy ions is neglectable.

>>>>

\section{Shock wave simulation}
We studied the evolution of trans-relativistic shocks and spectum of accelerated particles. We chose relastivistic shocks with low lorenz-factor because the maximum energy of produced cosmic rays increases with shock wave lorenz-factor, but the efficiency of acceleration decreases, because it is more difficult for particle to cross fast moving front many times \cite{Ellison2013}. So we assume that intermediate case of trans-relativistic shocks provides the most effective acceleration.

We developed the implicit particle-in-cell (PIC) code, presented in our previous paper \cite{Romansky2016}, based on the scheme suggested by Lapenta et al.~\cite{Lapenta2006} and improved for the relativistic case by Noguchi et al.\cite{Noguchi2007}.
Our code is fully three dimensional and parallelized with MPI technology, which is adapted for distributed computing and can be executed on a wide class of computers.

In simulation setup the homogeneous plasma flows in through the right boundary
and collides with the reflecting superconducting wall on the left boundary, causing the shock wave

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.86\textwidth]{fig/protons.eps} 
	\caption{Distribution of protons.}
	\label{protons}
\end{figure}
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.86\textwidth]{fig/electrons.eps} 
	\caption{Distribution of electrons.}
	\label{electrons}
\end{figure}

Simulations are one-dimensional and have following parameters: initial flow lorenz-factor $\gamma = 1.5$, number densities $n_e = 10^{-4} \rm{cm}^{-3}$, $n_p = 10^{-4} \rm{cm}^{-3}$ , temperature $5\cdot10^8 \rm{K}$, magnetic field $B = 10^{-4} \rm{G}$, the full size of the box $L = 2\cdot10^{12} \rm{cm}$, the number of cells $N=2\cdot10^4$. Electron mass is reduced to $m_e = \frac{m_p}{20}$. The full time of simulation is $T = 5000 {\omega_p}^{-1}$. $\theta$ is angle between flow velocity and magnetic field. We present results of particle spectrum in several simulations with different $\theta$.

One can see from Figures \ref{protons} and \ref{electrons} that spectrum of accelerated particles strongly depends on angle theta. If $\theta$ is less then critical value, defined by equation $c\cdot cos(\theta_{crit})=v_{shock}$, where all values are measured in upstream rest frame, particles can escape from the front and cross it several times to gain more energy. It is difficult to evaluate critical angle apriori, because we don't know compression relation of shock wave which will be created in simulation, but estimation for case of strong wave (comression relation $\sigma=4^{\circ}$) gets $\theta_{crit}=45$ in downstream frame. And it is consistent with results of our simulations.
